---
title: "Happiness Score Prediction"
author: "RABIN NEPAL (U00901360)"
date: "2023-11-26"
output: pdf_document
---

# Necessary Library Imports
```{r setup, include=FALSE}
# KNN
library(caret) 

# Ridge and Lasso Regression
# library(glmnet)

library(dplyr)

library(tidyr)

# for visualization
library(corrplot)
library(ggplot2)

# define some constants
# DATA_DIR = "/Users/rabinnepal/gitHub/Happiness-Prediction-Project/data"
DATA_DIR = "/home/ann/github/Happiness-Prediction-Project/data/"

# YEARS = list("2015","2016","2017","2018","2019","2020","2021","2022","2023")
YEARS = 2015:2023
YEARS = paste(YEARS) # convert num array to string array
TOTAL_YEARS = length(YEARS)
VARIABLE_NAMES = c()
RESPONSE = "Happiness_score"

# set working directory 
# setwd(DATA_DIR)
knitr::opts_knit$set(root.dir = DATA_DIR)

```


# Data Import and Cleanup
```{r}
# get filenames from data directory
files = list.files(DATA_DIR)
print(files)

# read csv files for world happiness data for each year 
# data_2015 = read.csv("2015.csv")
# data_2016 = read.csv("2016.csv")
# data_2017 = read.csv("2017.csv")
# data_2018 = read.csv("2018.csv")
# data_2019 = read.csv("2019.csv")
# data_2020 = read.csv("2020.csv")
# data_2021 = read.csv("2021.csv")
# data_2022 = read.csv("2022.csv")
# data_2023 = read.csv("2023.csv")

# read csv files for world happiness data for each year 
for(filename in files){
  year = substring(filename,1,4)
  # set name of variables dynamically as data_{year}
  variable_name = paste0("data_",year)
  VARIABLE_NAMES = c(VARIABLE_NAMES,variable_name)
  # print(variable_name)
  
  # read csv file
  year_data = read.csv(filename,stringsAsFactors = F)
   # year_data = read.csv(filename,stringsAsFactors = T)
  assign(variable_name,year_data)
  
  # see dimension of data for each year
  print(paste("Shape of data for ", year, "is :", dim(year_data)[1], "rows and", dim(year_data)[2], "columns."))
}


# see samples of each year data
# head(data_2015)
for(i in 1:TOTAL_YEARS){
  print(paste("A sample of data for year",2014+i,":"))
  dummy_year = get(VARIABLE_NAMES[i])
  glimpse(dummy_year)
}


# remove null values from dataset
for(i in 1:TOTAL_YEARS){
  dummy_year = get(VARIABLE_NAMES[i])
  if(any(is.na(dummy_year))){
    print("Found NA values and removing them...")
    dummy_year = na.omit(dummy_year)
    assign(VARIABLE_NAMES[i],dummy_year)
  }
}

```



# Data Exploration (All Years)
```{r}
# corrplot(corr = cor(data_2023[,-c(1,2)]),method = "color", outline = T,addCoef.col = "white", number.digits = 2, number.cex = 0.75)

# lets first see the columns for each year data
col_2015 = c(colnames(data_2015))
col_2016 = colnames(data_2016)
col_2017 = colnames(data_2017)
col_2018 = colnames(data_2018)
col_2019 = colnames(data_2019)
col_2020 = colnames(data_2020)
col_2021 = colnames(data_2021)
col_2022 = colnames(data_2022)
col_2023 = colnames(data_2023)

# lets find which columns are consistent among all these years
all_col = list(col_2015,col_2016,col_2017,col_2018,col_2019,col_2020,col_2021,col_2022,col_2023)
common_values = Reduce(intersect, all_col)
print(common_values) # gives 0 : means the columns have been named differently in the dataset

# After exploring the column names we can clearly see that the same variables are named differenty in the dataset for different years. (Ex: Country , Country.or.region , Country.name )

# lets make the column names uniform
column_names = c("Country", "Happiness.Score", "GDP.per.Capita", "Social.Support", "Life.Expectancy","Freedom", "Generosity", "Corruption")

# print(common_values) 
# gives all the values in column_names

# filter the values that are essential for the study (as mentioned on the world happiness report appendix)

for(i in 1:TOTAL_YEARS){
  dummy_year = get(VARIABLE_NAMES[i])
  # select according to columns names
  dummy_year = dummy_year[column_names]
  # sort by column name
  dummy_year = dummy_year[,order(colnames(dummy_year))]
  assign(VARIABLE_NAMES[i], dummy_year ) 
  }


# lets see which countries has held the No.1 rank in happiness over the years
happy_countries = c()
happy_countries_score = c()

sad_countries = c()
sad_countries_score = c()

for(i in 1:TOTAL_YEARS){
  dummy_year = get(VARIABLE_NAMES[i])
  happy_countries[i] = dummy_year$Country[1]
  happy_countries_score[i] = dummy_year$Happiness.Score[1]
  sad_countries[i] = dummy_year$Country[nrow(dummy_year)]
  sad_countries_score[i] = dummy_year$Happiness.Score[nrow(dummy_year)]
}

print(happy_countries)
print(sad_countries)

# see trend of happiness along the years 
barplot(happy_countries_score,names.arg=happy_countries,main = "Top Happiness Scores from 2015 to 2023", ylab = "Happiness Score",las = 2, col = 'coral')

# see trend of sadness along the years
barplot(sad_countries_score,names.arg=sad_countries,main = "Least Happiness Scores from 2015 to 2023", ylab = "Happiness Score",las = 2, col= "slateblue")


# a funcition to make the access of data easy
get_all_year_values = function(col_number){
  top_countries = c()
  top_countries_values = c()
  sad_countries = c()
  sad_countries_values = c()

  for(i in 1:TOTAL_YEARS){
    dummy_year = get(VARIABLE_NAMES[i])
    dummy_year = dummy_year[order(dummy_year[,col_number],decreasing = TRUE),]
    top_countries[i] = dummy_year$Country[1]
    top_countries_values[i] = dummy_year[1,col_number]
    sad_countries[i] = dummy_year$Country[nrow(dummy_year)]
    sad_countries_values[i] = dummy_year[nrow(dummy_year),col_number]
  }
  
  return(list(top_countries = top_countries,
              top_countries_values = as.numeric(top_countries_values),
              sad_countries = sad_countries,
              sad_countries_values = as.numeric(sad_countries_values)))
}

# see trend of corruption over the years
corruption_metric = get_all_year_values(1)
barplot(corruption_metric$top_countries_values,names.arg=corruption_metric$top_countries,main = "Highest Corruption Scores from 2015 to 2023", ylab = "Corruption Score",las = 2, col = 'coral')
barplot(corruption_metric$sad_countries_values,names.arg=corruption_metric$sad_countries,main = "Least Corruption Scores from 2015 to 2023", ylab = "Corruption Score",las = 2, col= "slateblue")


# see trend of freedom over the years
freedom_metric = get_all_year_values(3)
barplot(freedom_metric$top_countries_values,names.arg=freedom_metric$top_countries,main = "Highest Freedom Scores from 2015 to 2023", ylab = "Freedom Score",las = 2, col = 'coral')
barplot(freedom_metric$sad_countries_values,names.arg=freedom_metric$sad_countries,main = "Least Freedom Scores from 2015 to 2023", ylab = "Freedom Score",las = 2, col= "slateblue")


# see trend of gdp over the years
gdp_metric = get_all_year_values(4)
barplot(gdp_metric$top_countries_values,names.arg=gdp_metric$top_countries,main = "Highest GDP from 2015 to 2023", ylab = "GDP",las = 2, col = 'coral')
barplot(gdp_metric$sad_countries_values,names.arg=gdp_metric$sad_countries,main = "Least GDP from 2015 to 2023", ylab = "GDP",las = 2, col= "slateblue")


# see trend of genorosity over the years
generosity_metric = get_all_year_values(5)
barplot(generosity_metric$top_countries_values,names.arg=generosity_metric$top_countries,main = "Most Generous Countries from 2015 to 2023", ylab = "Freedom Score",las = 2, col = 'coral')
barplot(generosity_metric$sad_countries_values,names.arg=generosity_metric$sad_countries,main = "Least Generous Countries from 2015 to 2023", ylab = "Freedom Score",las = 2, col= "slateblue")


# see trend of life expectancy over the years
life_metric = get_all_year_values(6)
barplot(life_metric$top_countries_values,names.arg=life_metric$top_countries,main = "Country with Highest Life Expectancy from 2015 to 2023", ylab = "Freedom Score",las = 2, col = 'coral')
barplot(life_metric$sad_countries_values,names.arg=life_metric$sad_countries,main = "Country with Least Life Expectancy from 2015 to 2023", ylab = "Freedom Score",las = 2, col= "slateblue")


# see trend of Social support over the years
social_metric = get_all_year_values(7)
barplot(social_metric$top_countries_values,names.arg=social_metric$top_countries,main = "Country with Highest Social Support from 2015 to 2023", ylab = "Freedom Score",las = 2, col = 'coral')
barplot(social_metric$sad_countries_values,names.arg=social_metric$sad_countries,main = "Country with Lowest Social Support from 2015 to 2023", ylab = "Freedom Score",las = 2, col= "slateblue")


```



# Data Exploration (2023)
```{r}
# lets explore in details for a particular year
# summary of data
summary(data_2023)
#see relationship between all variables in data
pairs(data_2023[,-2],lower.panel=panel.smooth)

# see relation between happiness and different factors
cor(data_2023[,-2])

corrplot(corr = cor(data_2023[,-2]),method = "color", outline = T,addCoef.col = "white", number.digits = 2, number.cex = 0.75)

# top 5 happiest and unhappiest countries in 2023
top_5_happy = data_2023[ order(-data_2023$Happiness.Score),]
# top_5_happy = top_5_happy$Happiness.Score[1:5]

barplot(top_5_happy$Happiness.Score[1:5],names.arg=top_5_happy$Country[1:5], horiz=TRUE, main = "Top 5 happiest Countries in 2023", xlab = "Happiness Scores",las = 2, col = 'firebrick2')

barplot(top_5_happy$Happiness.Score[1:nrow()],names.arg=top_5_happy$Country[1:5], horiz=TRUE, add = TRUE,las = 2, col = 'firebrick2')

# # Add values at the end of each bar
# text(top_5_happy$Happiness.Score[1:5], 1:5, labels = top_5_happy$Happiness.Score[1:5], pos = 4, cex = 0.8, col = "black")



# top 5 generous and ungenerous countries in 2023
top_5_generous = data_2023[ order(-data_2023$Generosity),]

# top 5 healthiest and unhealthiest countries in 2023
top_5_healthiest = data_2023[ order(-data_2023$Life.Expectancy),]

# top 5 richest and poorest countries in 2023
top_5_richest = data_2023[ order(-data_2023$GDP.per.Capita),]

# top 5 free and restricted countries in 2023
top_5_free = data_2023[ order(-data_2023$Freedom),]


# top 5 corrupted and uncorrupted countries in 2023
top_5_corrupted = data_2023[ order(-data_2023$Corruption),]

```

# Data Clean-up for Regression

```{r}



```

# Train-test Split
```{r}
print(paste("Shape of data is :", dim(data_2023)[1], "rows and", dim(data_2023)[2], "columns."))



```

# Linear Regression Model

```{r}

```

# Lasso Regression Model

```{r}

```

# Ridge Regression Model

```{r}

```

# K-Nearest Neighbour Model

```{r}

```

# Comparison of Results

```{r}

```

## Accuracies

```{r}

```

| Models               |  Accuracy   |
|----------------------|-------------|
| Logistic Regression  |  79.38%     |
| LDA                  |  84.53%     |
| LDA (0.3)            |  84.53%     |
| Naive Bayes          |  83.50%     |
| KNN                  |  67.01%     |

## Plots

```{r}

```

## k-fold Cross Validation

```{r}

```



```{r cars}
summary(cars)
```



# References:
remove this link in final publication
* https://www.analyticsvidhya.com/blog/2022/01/different-types-of-regression-models/
* Dataset: 
https://medium.com/@TheSussex/is-happiness-free-bb9ef70d0bd5
https://www.kaggle.com/code/mshinde10/predicting-world-happiness
https://www.kaggle.com/code/kagleo123/world-happiness-record-eda-ml-regression
https://www.kaggle.com/datasets/unsdsn/world-happiness/code?datasetId=894&sortBy=dateRun
https://www.kaggle.com/datasets/unsdsn/world-happiness/code?datasetId=894&sortBy=dateRun


Dataset for 2020:
https://www.kaggle.com/datasets/londeen/world-happiness-report-2020

a
https://worldhappiness.report/data/
